---
title: "Early Extubation Resulting in Shorter Time to Post-Op Feeding"
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
#Load all the good stuff
library(tidyverse)
library(lubridate)
library(gt)
library(paletteer)
library(ggridges)

#Working directory for .RMD
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#Controlling figure output in markdown
knitr::opts_chunk$set(
#  fig.height =   
  fig.width = 6,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
 fig.align = "center",
  cache = FALSE
)
#Set Theme for ggplot2 - centers title and legend at bottom by default
theme_set(theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))

#Set Scientific notation output and decimal places for knitr
options(scipen = 999)
options(digits = 2)
```


```{r warning=FALSE, message=FALSE, echo = FALSE}
# Read in data
demographics <- read_csv("./data/demographics.csv")
info <- read_csv("./data/general_information.csv")
med_review <- read_csv("./data/medical_review.csv")
op2 <- read_csv("./data/operative_information.csv")
postop <- read_csv("./data/postoperative_information.csv")
```

```{r warning=FALSE, message=FALSE, echo = FALSE}
# Select first round of desired variables from each data set
demographics <- demographics %>%
  dplyr::select(BLIND_ID, WEIGHT, SEX, RACE, RACE_W, RACE_B, other_race, ETHNICITY)

info <- info %>%
  dplyr::select(BLIND_ID, BLIND_SITE_ID)

med_review <- med_review %>%
  dplyr::select(BLIND_ID, PROC_AGE, DIAG, CHROM_22, CHROM_21, CHROM_18, CHROM_CRI, CHROM_OTH)

op2 <- op2 %>%
  dplyr::select(BLIND_ID, OP_W, INTUB_Y_N, intub_d_day, INTUB_T, INTUB_OR_T, BYPASS, NO_XCLAMP_T, XCLAMP_T, CPB_T, EXTUB_OR_YN, EXTUB_OR_T, EXIT_T)

postop <- postop %>%
  dplyr::select(BLIND_ID, icu_d_day, ICU_T, EXTUB30_ICU_YN, EXTUB30_ICU_T, EXTUB6_ICU_YN, EXTUB6_T, extub_d_day, EXTUB_T, RE_INTUB, re_intub_d_day, RE_INTUB_T, re_extub_d_day, RE_EXTUB_T, ENTERAL, enteral_d_initiate_day, ENTERAL_T_INITIATE, ORAL, enteral_d_oral_initiate_day, ENTERAL_T_ORAL_INITIATE, TUBE)
```

```{r warning=FALSE, message=FALSE, echo = FALSE}
# Store data sets as a list
data_list <- list(demographics, info, med_review, op2, postop)

# Join data sets
df_full <- data_list %>%
  reduce(inner_join, by = "BLIND_ID") %>%
  janitor::clean_names() %>% # Clean up variable names
  rename(race_known = race) # Rename to prevent confusion

# Make a single race variable
df_full <- df_full %>%
  mutate(race = case_when(
    race_w == "Yes" ~ "White",
    race_b == "Yes" ~ "Black/AA",
    other_race == "Yes" ~ "Another race",
  )) %>%
    mutate(race = ifelse(is.na(race), "Unknown", race)) 

# Make y/n chromosomal anomaly variable
df_full <- df_full %>%
  mutate(chrom_anomaly = case_when(
    chrom_22 == "Yes" ~ "Yes",
    chrom_21 == "Yes" ~ "Yes",
    chrom_18 == "Yes" ~ "Yes",
    chrom_cri == "Yes" ~ "Yes",
    chrom_oth == "Yes" ~ "Yes",
    TRUE ~ "No"
  ))

# Create early extubation (y/n) variable
df_full <- df_full %>%
  mutate(early_extub = case_when(
    extub_or_yn == "Yes" ~ "Yes", # if extubated in the OR
    extub30_icu_yn == "Yes" ~ "Yes", # if extubated within 30 minutes of ICU arrival
    extub6_icu_yn == "Yes" ~ "Yes", # if extubated within 6 hours
    TRUE ~ "No"
  ))

# Check if any infants in the early extubation group were reintubated
d <- df_full %>%
  dplyr::select(early_extub, re_intub) %>%
  filter(early_extub == "Yes")

# Modify early extubation variable to be "No" if infants were reintubated
df_full <- df_full %>%
  mutate(early_extub = case_when(
    re_intub == "Yes" ~ "No",
    TRUE ~ early_extub
  ))

# Explore NAs in days to initial extubation variable
d <- df_full %>%
  dplyr::select(extub_d_day, early_extub) %>%
  filter(early_extub == "Yes")
# If early extubation is "Yes", then the days to extubation variable is NA
# which essentially means that they were all extubated on the day of surgery.
# Therefore...

# Create new days to extubation variable where these NAs are changed to 0 days
df_full <- df_full %>%
  mutate(extub_days = case_when(
    early_extub == "Yes" ~ 0,
    TRUE ~ extub_d_day
  ))

# For those who were reintubated, replace value with days to final (second) extubation
df_full <- df_full %>%
  mutate(
    extub_days =
      ifelse(re_intub == "Yes", re_extub_d_day, extub_days)
  )

# Check
d <- df_full %>%
  dplyr::select(extub_d_day, early_extub, extub_days, re_intub, re_extub_d_day) %>%
  filter(re_intub == "Yes")

# Create variable for days to first oral feeding after extubation
df_full <- df_full %>%
  mutate(oral_postextub = (enteral_d_oral_initiate_day - extub_days))

# Explore NAs in days to first oral feed variable
d <- df_full %>%
  dplyr::select(oral, enteral_d_oral_initiate_day, oral_postextub, extub_d_day, early_extub, extub_days, re_intub, re_extub_d_day) %>%
  filter(is.na(oral_postextub))
# There are 36 NAs due to those infants not starting oral feeding before ICU discharge

# Remove NAs (didn't feed orally before ICU discharge) and 9 cases with implausible values for time to first oral feed (negative value)
df_full <- df_full %>%
  filter(oral_postextub >= 0)
# n = 277

# Create variable with max cross clamp time
df_full <- df_full %>%
  mutate(cc_time = pmax(no_xclamp_t, xclamp_t, na.rm = TRUE))

# Modify cardiopulmonary bypass (CPB) time variable to be 0 if there was no CPB used
df_full <- df_full %>%
  mutate(cpb_time = case_when(
    bypass == "No" ~ 0,
    TRUE ~ cpb_t
  ))

# Rename levels
df_full <- df_full %>%
  mutate(
    ethnicity = factor(ethnicity), # ethnicity
    hispanic = fct_recode(ethnicity,
      "Yes" = "Hispanic or Latino/Latina",
      "No" = "Not Hispanic or Latino/Latina",
      "Unknown" = "Unknown"
    )
  ) %>%
  mutate(
    diag = factor(diag), # diagnosis
    diagnosis = fct_recode(diag,
      "CoA" = "Coarctation of Aorta Repair",
      "ToF" = "Tetralogy of Fallot Repair"
    )
  )

# Select variables for final data set
df_final <- df_full %>%
  dplyr::select(blind_id, blind_site_id, sex, race, hispanic, diagnosis, chrom_anomaly, proc_age, op_w, cpb_time, cc_time, early_extub, extub_days, enteral_d_oral_initiate_day, enteral_t_oral_initiate, oral_postextub)

# Change variables to factors
# Create vector with columns to change to factors
cols <- c("blind_id", "blind_site_id", "sex", "race", "hispanic", "diagnosis", "chrom_anomaly", "early_extub")

```

```{r warning= FALSE, echo = FALSE}

#average age of children - days
age_avg_d <- (mean(df_final$proc_age))

#average age of children - years
age_avg_y <- (mean(df_final$proc_age) / 365)
```


## Summary Statistics
<br>
<br>
We will begin the analysis by reviewing the summary statistics on the __df_final__ dataset (n = 322) from the Pediatric Heart Network in order to identify if early extubation has any association with shorter time to the first postoperative oral feeding, and whether this differed based on the diagnosis group, tetralogy of fallout or coarctation of the aorta.
<br>
<br>

```{r warning= FALSE, echo = FALSE}

#Convert factors to numeric for summary statistics
sum_stats <- df_final %>% 
  mutate(diagnosis = ifelse(df_final$diagnosis == "CoA", 0 ,1),
         early_extub = ifelse(df_final$early_extub == "No", 0, 1))


# Summary stats
sum_stats_final <- sum_stats %>% 
  group_by(blind_id) %>% 
  summarise(
    "# Days to Post-op Feeding" = unique(enteral_d_oral_initiate_day),
    "% Early Extubated (Yes)" = (mean(early_extub) * 100),
    "% Diagnosis (Tetralogy)" = (mean(diagnosis) * 100)) %>% 
    ungroup() %>% 
      pivot_longer(
        cols = contains(c("#", "%")),
          names_to = "Category",
          values_to = "Value"
      )

#gt() output
sum_stats_final %>% 
  group_by(Category) %>% 
  summarise(
    across(
      contains(c("Value")),
      list(Mean = mean, SD = sd), na.rm = TRUE,
      .names = "{.fn}")) %>% 
    gt() %>% 
    fmt_number(columns = 3, decimals = 2) %>% 
    fmt_number(columns = 2, decimals = 2) %>% 
    tab_header("Summary Statistics")
```
<br>
<br>
Based on the final dataset, the average age of the individual undergoing surgery was `r age_avg_d` days and `r age_avg_y` years