---
title: "*Early Extubation Resulting in Shorter Time to Post-Op Feeding*"
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
#Load all the good stuff
library(tidyverse)
library(lubridate)
library(gt)
library(plotly)
library(paletteer)
library(ggridges)

#Working directory for .RMD
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#Controlling figure output in markdown
knitr::opts_chunk$set(
#  fig.height =   
  fig.width = 6,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
 fig.align = "center",
  cache = FALSE
)
#Set Theme for ggplot2 - centers title and legend at bottom by default
theme_set(theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))

#Set Scientific notation output and decimal places for knitr
options(scipen = 999)
options(digits = 4)
```

```{r warning=FALSE, message=FALSE, echo = FALSE}
#Read data for this research question
df_rq2 <- read_csv("./source/df_final.csv")
```


```{r warning= FALSE, echo = FALSE}

#average age of children - days
age_avg_d <- round(mean(df_rq2$proc_age), 2)

#average age of children - years
age_avg_y <- round(mean(df_rq2$proc_age) / 365, 2)
```

<br>

## Exploratory Data Analysis

We will begin the exploratory analysis by reviewing the summary statistics on the _df_final_ dataset (n = 322) from the Pediatric Heart Network in order to identify if early extubation has any association with shorter time to the first postoperative oral feeding, and whether this differed based on the diagnosis group, Tetralogy of Fallot or coarctation of the aorta.
<br>
<br>

```{r warning= FALSE, echo = FALSE}

#Convert factors to numeric for summary statistics
sum_stats <- df_rq2 %>% 
  mutate(diagnosis = ifelse(df_rq2$diagnosis == "CoA", 0 ,1),
         early_extub = ifelse(df_rq2$early_extub == "No", 0, 1))


# Summary stats
sum_stats_final <- sum_stats %>% 
  group_by(blind_id) %>% 
  summarise(
    "# Days to Post-op Feeding" = unique(enteral_d_oral_initiate_day),
    "% Early Extubated (Yes)" = (mean(early_extub) * 100),
    "% Diagnosis (Tetralogy)" = (mean(diagnosis) * 100)) %>% 
    ungroup() %>% 
      pivot_longer(
        cols = contains(c("#", "%")),
          names_to = "Category",
          values_to = "Value"
      )

#gt() output
sum_stats_final %>% 
  group_by(Category) %>% 
  summarise(
    across(
      contains(c("Value")),
      list(Mean = mean, SD = sd), na.rm = TRUE,
      .names = "{.fn}")) %>% 
    gt() %>% 
    fmt_number(columns = 3, decimals = 2) %>% 
    fmt_number(columns = 2, decimals = 2) %>% 
    tab_header("Summary Statistics") 
```
<br>
<br>
Based on the final dataset, the average age of the individual undergoing surgery was **`r age_avg_d`** days, or **`r age_avg_y`** years. The data was relatively clean and the final dataset resulted in no missingness. The summary statistics show that the average number of days to first postoperative oral feeding is **1.90** days, with a standard deviation of **4.44** days. The percent of surgical procedures that had a diagnosis of Tetralogy of Fallot is **62.82**% and coarctation of the aorta is **37.18**%. The percent of surgical procedures that experienced early extubation (extubation within 6 hours of surgery) is **53.79**% and those that did not experience early extubation is **46.21**%.
<br>
<br>

## Visualization

```{r results="asis", warning = FALSE, message = FALSE, echo = FALSE}
#DF for visualizing extubation and days until first post-op feeding
rq2_figure <- df_rq2 %>% 
  group_by(blind_id, enteral_d_oral_initiate_day, diagnosis, early_extub) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(
    diagnosis = recode(diagnosis, 'ToF' = 'Tetralogy of Fallot', 'CoA' = 'Coarctation of the Aorta')
  ) %>% 
  ggplot(aes(x = enteral_d_oral_initiate_day, y = n, fill = early_extub)) + 
  geom_bar(stat = "identity",  position = "stack") + 
  labs(y = '# Patients Receiving First Post-Op Oral Feeding',
       x = '# of Days Until First Post-Op Oral Feeding') +
  viridis::scale_fill_viridis(
    name = 'Early Extubation',
    discrete = T
  )

#generate ggplotly for first figure
rq2.1 <- ggplotly(rq2_figure)

#DF for visualizing extubation and days until first post-op feeding by diagnosis group
rq2_figure_diag <- df_rq2 %>% 
  group_by(blind_id, enteral_d_oral_initiate_day, diagnosis, early_extub) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(
    diagnosis = recode(diagnosis, 'ToF' = 'Tetralogy of Fallot', 'CoA' = 'Coarctation of the Aorta')
  ) %>% 
  ggplot(aes(x = enteral_d_oral_initiate_day, y = n, fill = early_extub)) + 
  geom_bar(stat = "identity",  position = "stack") + 
  labs(y = '# Patients Receiving First Post-Op Oral Feeding',
       x = '# of Days Until First Post-Op Oral Feeding') +
  theme(plot.title.position = 'plot', 
        plot.title = element_text(hjust = 15)) +
  facet_grid(~diagnosis) + 
  viridis::scale_fill_viridis(
    name = 'Early Extubation',
    discrete = T
  )

#generate ggplotly for second figure
rq2.2 <- ggplotly(rq2_figure_diag)

#DF for visualizing days until first post-op feeding by diagnosis group and extubation status as facets
rq2_figure_diag_ext <- df_rq2 %>% 
  group_by(blind_id, enteral_d_oral_initiate_day, diagnosis, early_extub) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(
    diagnosis = recode(diagnosis, 'ToF' = 'Tetralogy of Fallot', 'CoA' = 'Coarctation of the Aorta')
  ) %>% 
  ggplot(aes(x = enteral_d_oral_initiate_day, y = n)) + 
  geom_bar(stat = "identity",  position = "stack") + 
  labs(y = '# Patients Receiving First Post-Op Oral Feeding',
       x = '# of Days Until First Post-Op Oral Feeding') +
  facet_grid(early_extub ~ diagnosis) + 
  viridis::scale_fill_viridis(
    name = 'Early Extubation',
    discrete = T
  )

#generate ggplotly for third figure
rq2.3 <- ggplotly(rq2_figure_diag_ext)
```
<br>

#### **Frequency of Days until Post-Op Oral Feeding By Early Extubation Status**

`r rq2.1`

<br>

#### **Frequency of Days until Post-Op Oral Feeding By Early Extubation Status Grouped by Diagnosis**

`r rq2.2`

<br>

#### **Frequency of Days until Post-Op Oral Feeding Grouped by Early Extubation Status and Diagnosis**

`r rq2.3`

<br>


## Statistcal Analysis Results
<br>

Based on the visualizations, we are able to see that a majority of patients receiving first post-operative oral feeding occurred within 1 - 2 days, with those receiving early extubation resulting in shorter times. The diagnosis of Tetralogy of Fallot also demonstrates shorter times to first post-operative oral feeding when separating the diagnosis in separate grids.

We decided to proceed with testing our hypothesis on early extubation or diagnosis having an association with shorter time to post-operative oral feeding. We fit a Poisson regression model since the outcome or response variable is looking to predict the count or rate of outcome. Additionally, we ran a residuals versus fitted model and checked for overdispersion.


$$ log(Time \ to \ first \ postoperative \ oral \ feeding \ (days)) \ = \ a + \beta_1 \ early \ extubation + \beta_2 \ diagnosis$$
```{r results="asis", warning = FALSE, message = FALSE, echo = FALSE}
# Add color for kable titles
cat("
<style>
caption {
      color: black;
      font-weight: bold;
      font-size: 1.0em;
    }
</style>
")
```
```{r warning= FALSE, echo = FALSE, message = FALSE}

#Linear model to determine associaton
poiss_model <- glm(enteral_d_oral_initiate_day ~ early_extub +
                     diagnosis, data = df_rq2,
                   family = poisson(link = "log")) 

#Plot for only residuals vs fitted
plot(poiss_model, which = 1)


#Model output with knitr
poiss_model %>%
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(caption = "Poisson Model Results")


#Risk ratio
exp(coef(poiss_model)) %>% 
  broom::tidy() %>% 
  knitr::kable(caption = "Risk Ratio")

#Confidence interval
exp(confint(poiss_model)) %>% 
  knitr::kable(caption = "95% Confidence Interval")

#Checking for overdispersion and model fit
pchisq(poiss_model$deviance, poiss_model$df.residual, lower.tail = FALSE) %>% 
  broom::tidy() %>% 
  knitr::kable(caption = "Check for Overdispersion")

```
<br>

## Conclusion
<br>
Based on the model results, we can suggest that early extubation is associated with shorter time to postoperative oral feeding while also based on diagnosis type. Additionally, the diagnosis group also display statistical significance based on reducing the time to the first postoperative oral feed. This conclusion was drawn from the _p-value_ for both outcomes falling below the threshold of significance (< 0.05) along with a 95% confidence interval not including 1 in either outcome. 

Adjusted for diagnosis, we estimate that infants who received early extubation had a 69% shorter time to postoperative oral feeding (95% CI: 75%-63% shorter time, p<0.001). Additionally, holding early extubation status constant, we estimate that infants undergoing surgery for Tetralogy of Fallot experienced a 28% shorter time to postoperative oral feeding than those undergoing repair for coarctation of the aorta (95% CI: 40%-15% shorter time, p<0.001). 

